---
title: "API Call"
author: "Cherry Pham"
date: "2023-06-20"
output: html_document
---

This Markdown file is created with the purpose of obtaining sensor data from Quant-AQ through API calls.

There is no prerequisites for using this file. If you are just looking to quickly retrieve data, you can simply run this file through the command `Ctrl + Alt + R`. The result of running this file is a data folder containing data of MOD-PM sensors as of February 2023. If you wish to modify the data to be retrieved, please read the data guide below.

## DATA GUIDE

[write instructions here]

[sorry if i forgot to do this at the end, the code below is quite well documented though so good luck]

## PACKAGES AND INITIAL SETTINGS

```{r}
# Check and install required packages if necessary
packages <- c("httr", "jsonlite", "purrr", "dplyr")
install.packages(packages[!sapply(packages, requireNamespace, quietly = TRUE)])

# Load required packages
invisible(sapply(packages, library, character.only = TRUE))

# Set options
knitr::opts_chunk$set(echo=FALSE, message=FALSE)
```

## DEFINE VARIABLES

```{r}
base_url <- "https://api.quant-aq.com/device-api/v1/devices/"
api_key <- "VUOLNUDTM70QZ7Z9G0Z78XA7"

# Define a time range to loop over until I find a better way
time_range <- seq(as.Date("2022-07-01"), as.Date("2022-09-30"), by = "day")
time_range <- format(time_range, "%Y-%m-%d")

# Define a list of device IDs
device_ids <- c(
  "MOD-00024",
  "MOD-00025",
  "MOD-00026",
  "MOD-00027",
  "MOD-00028",
  "MOD-PM-00141",
  "MOD-PM-00211",
  "MOD-PM-00212",
  "MOD-PM-00213",
  "MOD-PM-00214",
  "MOD-PM-00216",
  "MOD-PM-00217",
  "MOD-PM-00221",
  "MOD-PM-00222",
  "MOD-PM-00224",
  "MOD-PM-00226",
  "MOD-PM-00230",
  "MOD-PM-00231"
)
```

## RETTRIEVING DATA INTO FOLDERS

```{r}
# Create folder if not initiated
data_folder_path <- "./data/MOD-PM/"

# Check if the folder exists
if (!file.exists(data_folder_path)) {
  # Create the folder if it doesn't exist
  dir.create(data_folder_path, recursive = TRUE)
}

# Function to process a single device ID
process_device_id <- function(device_id, date) {
  # Construct device URL for API call
  device_url <- paste0(device_id, "/data-by-date/raw/", date, "/")
  
  # Send GET request
  raw_data <- GET(
    url = paste0(base_url, device_url),
    authenticate(api_key, "", type = "basic"),
    encoding = "UTF-8"
  )
  
  data <- content(raw_data, as = "text", encoding = "UTF-8")
  parsed_data <- fromJSON(data)
  
  # Check if the data is empty or has zero rows
  if (is.null(parsed_data) || is.null(parsed_data$data) || length(parsed_data$data) == 0) {
    # Skip processing if the data is empty
    return()
  }

  # Extract relevant data
  df <- as.data.frame(parsed_data$data)
  
  device_folder_path <- paste0("./data/MOD-PM/", device_id)
  if (!file.exists(device_folder_path)) {
    dir.create(device_folder_path, recursive = TRUE)
  }

  # Save data to CSV file named after the device ID
  csv_file_path <- paste0("./data/MOD-PM/", device_id, "/", device_id, "-", date, ".csv")
  
  # Try writing the CSV file and handle errors
  tryCatch(
    {
      # Create the device - specific folder if it doesn't exist
      write.csv(df, file = csv_file_path, row.names = FALSE, fileEncoding = "UTF-8")
    },
    error = function(e) {
      print(paste0(device_id, " ", date))
      print(paste("Error writing CSV file for", device_id, "on", date))
    }
  )
}

# Process each device ID over time range
for (device_id in device_ids) {
  for (date in time_range) {
    process_device_id(device_id, date)
  }
}
```
