---
title: "data_viz"
author: "Cherry Pham"
date: "2023-07-06"
output: html_document
---

[Markdown file docstrings]

## INITAL SETTINGS

```{r}
# Check and install required packages if necessary
packages <- c(
  "lubridate",   # Package for working with dates and times
  "data.table",  # Package for efficient data manipulation and processing
  "dplyr",       # Package for data manipulation and transformation
  "openair",     # Package for analyzing air pollution data
  "openairmaps", # Package for accessing open air pollution data
  "stringr",     # Package for string manipulation
  "baseline",    # Package for baseline modeling and adjustment
  "purrr",       # Package for functional programming
  "tidyverse",   # Meta-package for data science
  "maps"         # Package for creating maps
)
install.packages(packages[!sapply(packages, requireNamespace, quietly = TRUE)])

# Load required packages for data manipulation and analysis
invisible(sapply(packages, library, character.only = TRUE))

# Set options
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

# Suppress Warnings (This is a me thing, comment this session out to see the conflicting functions)
suppressPackageStartupMessages(library(maps))
```

## FETCHING DATA TO IMPORT

To skip this session, go to the data folder branch of [this repository](https://github.com/airpartners/ambient-summer23/tree/datafolder) to download a pre-made data folder. If you want to modify data/download data for yourself, refer to the below instructions within this session.

### Quant-AQ Data

Refer to the Markdown file API_call.Rmd in this folder. The data should be in ./data/Quant-AQ/MOD-PM/ if nothing goes wrong.

### Meteorology Data

The meteorology data used in some analysis below are taken from the [Iowa Environmental Mesonet](https://mesonet.agron.iastate.edu/request/download.phtml?network=MA_ASOS).

Below are the downloading instructions taken from the [EB Gibson Park Analysis](https://github.com/scott-hersey/EB_AQ_Network/blob/master/Gibson%20Park%20Analysis.Rmd).

Select the following station: \* [BOS] BOSTON/LOGAN INTL

Select the following variables:

-   Temperature (C)

-   Wind Direction

-   Wind Speed [mph]

Select the date range: `7/1/2022 to 2/28/2022` (this dataset is not inclusive of the last date)

Select this timezone: `America/New_York`

Use the following download options: `comma delimited` ,

`no latitude/longitude vectors` ,

`no elevation` ,

`represent missing data with blank string` ,

`denote trace with blank string`

## IMPORT DATA AND PRE-PROCESSING

### Quant-AQ Data

```{r}
# Read in csv files from data folder
device_ids <- c(
  "MOD-00024",
  "MOD-00025",
  "MOD-00026",
  "MOD-00027",
  "MOD-00028",
  "MOD-PM-00141",
  "MOD-PM-00211",
  "MOD-PM-00212",
  "MOD-PM-00213",
  "MOD-PM-00214",
  "MOD-PM-00216",
  "MOD-PM-00217",
  "MOD-PM-00221",
  "MOD-PM-00222",
  "MOD-PM-00224",
  "MOD-PM-00226",
  "MOD-PM-00230",
  "MOD-PM-00231"
)

modPM <- suppressMessages({
  data <- device_ids %>%
    map_dfr(function(sensor_id) {
      paths <- sprintf(
        "./data/Quant-AQ/MOD-PM/broken-down/%s/%s-2022-%02d.csv",
        sensor_id, sensor_id, 7:9
      )
      map_dfr(paths, function(path) {
        suppressWarnings(
          tryCatch(
            read.csv(path),
            error = function(e) NULL
          )
        )
      })
    })
  data
})
```

```{r}
# Clean up Quant-AQ data
modPM <- modPM %>%
  setnames(old = c(
    "timestamp.x",
    "sn.x",
    "opc.bin0",
    "opc.bin1",
    "opc.bin2",
    "opc.bin3",
    "opc.bin4",
    "opc.bin5",
    "opc.bin6",
    "opc.bin7",
    "opc.bin8",
    "opc.bin9",
    "opc.bin10",
    "opc.bin11",
    "opc.bin12",
    "opc.bin13",
    "opc.bin14",
    "opc.bin15",
    "opc.bin16",
    "opc.bin17",
    "opc.bin18",
    "opc.bin19",
    "opc.bin20",
    "opc.bin21",
    "opc.bin22",
    "opc.bin23",
    "opc.pm1",
    "opc.pm25",
    "opc.pm10",
    "neph.pm1",
    "neph.pm25",
    "neph.pm10",
    "neph.cscat",
    "gases.co.we",
    "gases.co.ae",
    "gases.co.diff",
    "gases.no.we",
    "gases.no.ae",
    "gases.no.diff",
    "gases.no2.we",
    "gases.no2.ae",
    "gases.no2.diff",
    "gases.o3.we",
    "gases.o3.ae",
    "gases.o3.diff",
    "modelpmpm1",
    "modelpmpm25",
    "modelpmpm10"
  ),
  new = c(
    "timestamp",
    "sn",
    "bin0",
    "bin1",
    "bin2",
    "bin3",
    "bin4",
    "bin5",
    "bin6",
    "bin7",
    "bin8",
    "bin9",
    "bin10",
    "bin11",
    "bin12",
    "bin13",
    "bin14",
    "bin15",
    "bin16",
    "bin17",
    "bin18",
    "bin19",
    "bin20",
    "bin21",
    "bin22",
    "bin23",
    "opcn3_pm1",
    "opcn3_pm25",
    "opcn3_pm10",
    "pm1_env",
    "pm25_env",
    "pm10_env",
    "neph_csat",
    "co_we",
    "co_ae",
    "co_diff",
    "no_we",
    "no_ae",
    "no_diff",
    "no2_we",
    "no2_ae",
    "no2_diff",
    "o3_we",
    "o3_ae",
    "ox_diff",
    "pm1_model_id",
    "pm25_model_id",
    "pm10_model_id"
  ))
```

```{r}
# Combine pm1 bins
modPM$pm1num <- modPM$bin0 + modPM$bin1 + modPM$bin2

# Add a formatted date column
modPM$date <- as.POSIXct(strptime(modPM$timestamp_local, format = "%Y-%m-%d %H:%M:%S", tz = "America/New_York"))

# Delete unnecessary columns
modPM <- subset(modPM, select = -c(
  url.x, url.y, sn.y, timestamp.y, geo.lat, geo.lon, met.rh, met.temp, met.wd, met.ws, geolat, geolon, metrh, geo.xlat,
  geo.xlon, met.xrh, met.xtemp, nephbin0, nephbin1, nephbin2, nephbin3, nephbin4, nephbin5, nephpm1, nephpm10,
  nephpm25, opcbin0, opcbin1, opcbin10, opcbin11, opcbin12, opcbin13, opcbin14, opcbin15, opcbin16, opcbin17, opcbin18,
  opcbin19, opcbin2, opcbin20, opcbin21, opcbin22, opcbin23, opcbin3, opcbin4, opcbin5, opcbin6, opcbin7, opcbin8,
  opcbin9, opcpm1, opcpm10, opcpm25, opcrh, opctemp, geo.ylat, geo.ylon, met.yrh, met.ytemp
))
print(modPM)
```

### Meteorology Data

```{r}
# Import meteorology file
metdata <- suppressMessages({
  list.files(path = "./data/Metdata/",
             pattern = "*.csv", 
             full.names = TRUE) %>%
    purrr::map_dfr(read.csv)
})

print(metdata)
# Setting datetime, using correct timezone on East Coast Local time
metdata$date <- as.POSIXct(metdata$valid, format = "%Y-%m-%d %H:%M", tz = "America/New_York")

metdata <- metdata %>%
  # if only wind speed and wind direction matter
  setnames(old = c("drct", "sped", "valid"), new = c("wd", "ws", "original_met_time")) %>% #rename
  na.omit("date") %>% #if any dates are NA, the following function won't work
  complete(date = seq(from = min(date), to= max(date),  by = "1 min")) %>%#make 1 minute intervals
  fill(c("wd", "ws", "tmpc")) #fill those new 1 minute interval rows

metdata$ws <- metdata$ws * (1609/3600) #converting to m/s, 1609 meters per mile, 3600 seconds per hr
metdata[,c("station")] <- list(NULL) #getting rid of unnecessary variables
metdata <- unique(metdata) #remove duplicates
print(metdata)
```

### Merging Data-sets

```{r}
# Time sync and merge met and pollution data
modPM$modPM_date_1min <- lubridate::round_date(modPM$date, unit = "minute")
metdata$met_date_1min <- lubridate::round_date(metdata$date, unit = "minute")

# Merge the two datasets
modPM_met <- dplyr::left_join(modPM, metdata, by = c("modPM_date_1min" = "met_date_1min"))
modPM_met$date <- modPM_met$date.x

# Delete unnecessary columns
modPM_met <- subset(modPM_met, select = -c(date.x, date.y))
```

### Sanity Checks

```{r}
print(modPM_met)

openair::timeVariation(modPM_met, pollutant = c("ws", "wd", "tmpc"), normalise = TRUE)
```

## GRAPHING

### Graph 1

Create a time series plot in r comparing bus/train arrival times to pollutant concentration, filtered for wind speed and/or direction

```{r}
# # Merge the datasets based on the timestamp column
# merged_data <- merge(arrival_times, pollutant_data, by = "timestamp", all = TRUE)
# 
# # Filter the merged data based on wind speed and/or direction
# filtered_data <- merged_data %>%
#   filter(wind_speed > threshold_value)
# 
# # Time series comparing bus/train arrival times to pollutant concentration
# ggplot(filtered_data, aes(x = timestamp)) +
#   geom_line(aes(y = arrival_times, color = "Arrival Times")) +
#   geom_line(aes(y = pollutant_concentration, color = "Pollutant Concentration")) +
#   labs(x = "Timestamp", y = "Values") +
#   scale_color_manual(values = c("Arrival Times" = "blue", "Pollutant Concentration" = "red")) +
#   theme_minimal()
```

### Graph 2

Create bar graphs with each bar representing the pollutants from a single sensor (stacking 3 colors: pink for PM1, yellow for PM2.5, blue for PM10)

```{r}
library(ggplot2)  # Load the ggplot2 package

# Create a subset of the data for PM1, PM2.5, and PM10
pm_subset <- modPM[, c("sn", "pm1", "pm25", "pm10")]

# Reshape the data into long format
pm_subset_long <- tidyr::pivot_longer(pm_subset, cols = c(pm1, pm25, pm10), names_to = "pollutant", values_to = "value")

# Define the colors for each pollutant
colors <- c("pm1" = "orange", "pm25" = "yellow", "pm10" = "blue")

# Create the bar graph
ggplot(pm_subset_long, aes(x = sn, y = value, fill = pollutant)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = colors) +
  labs(x = "Sensor ID", y = "Pollutant Value") +
  ggtitle("Pollutant Levels by Sensor") +
  theme_minimal()
```

### Graph 3

Create a bar graph, where vertical axis is number of days in each month, horizontal axis is months, and the bars are stacked for number of days with good, medium and bad air quality for each sensor.

```{r}
library(ggplot2)  # Load the ggplot2 package

# Define air quality thresholds
good_threshold <- 10  # Placeholder value for good air quality
medium_threshold <- 20  # Placeholder value for medium air quality

# Calculate the number of days in each category for each sensor in each month
air_quality_summary <- modPM %>%
  mutate(month = lubridate::month(date),
         air_quality_category = case_when(
           pmsum <= good_threshold ~ "Good",
           pmsum <= medium_threshold ~ "Medium",
           TRUE ~ "Bad"
         )) %>%
  group_by(sn, month, air_quality_category) %>%
  summarise(num_days = n()) %>%
  ungroup()

# Set the order of the air_quality_category factor levels
air_quality_summary$air_quality_category <- factor(air_quality_summary$air_quality_category, levels = c("Bad", "Medium", "Good"))

# Create the stacked bar graph
ggplot(air_quality_summary, aes(x = month, y = num_days, fill = air_quality_category)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Good" = "green", "Medium" = "orange", "Bad" = "red")) +
  labs(x = "Month", y = "Number of Days", fill = "Air Quality") +
  ggtitle("Air Quality Summary by Month and Sensor") +
  theme_minimal() +
  facet_grid(sn ~ ., scales = "free_x", labeller = labeller(sn = sn_labels)) +
  geom_text(aes(label = num_days), position = position_stack(vjust = 0.5), color = "white")
```

### Graph 4

Create a map with the average polar plots w/ of each pollution values (PM1, PM2.5, PM10) over certain periods of time (all time, monthly, daily, hourly)

```{r}
# Sample data (replace with your actual data)
pollution_data <- data.frame(
  Longitude = c(-74.0060, -73.9352, -73.9919, -73.9962, -73.9812),
  Latitude = c(40.7128, 40.7309, 40.7589, 40.7488, 40.7549),
  PM1 = c(5.2, 6.3, 4.8, 5.9, 6.1),
  PM2.5 = c(10.1, 12.5, 9.8, 11.2, 12.4),
  PM10 = c(15.6, 18.9, 14.2, 16.8, 17.9),
  DateTime = as.POSIXct(c("2023-01-01 12:00:00", "2023-01-01 12:00:00", "2023-01-01 12:00:00",
                          "2023-01-01 12:00:00", "2023-01-01 12:00:00"))
)

# Create the average polar plots
avg_polar_plots <- polarPlot(pollution_data, pollutant = c("PM1", "PM2.5", "PM10"),
                             type = "average", res = 1, start = "all",
                             polar = TRUE, ci = FALSE, nsector = 12)

# Plot the polar plots on a map
plot(avg_polar_plots, map = TRUE)
```

### Graph 5

Create a map of for the average of each pollution values (PM1, PM2.5, PM10) over certain periods of time (all time, monthly, daily, hourly)

```{r}
# Sample data (replace with your actual data)
pollution_data <- data.frame(
  Longitude = c(-74.0060, -73.9352, -73.9919, -73.9962, -73.9812),
  Latitude = c(40.7128, 40.7309, 40.7589, 40.7488, 40.7549),
  PM1 = c(5.2, 6.3, 4.8, 5.9, 6.1),
  PM2.5 = c(10.1, 12.5, 9.8, 11.2, 12.4),
  PM10 = c(15.6, 18.9, 14.2, 16.8, 17.9),
  Date = as.Date(c("2023-01-01", "2023-01-01", "2023-01-01", "2023-01-01", "2023-01-01"))
)

# Define the map boundaries
map_limits <- map("world", "USA", xlim = c(-125, -66), ylim = c(24, 50), plot = FALSE)

# Function to plot the map with average pollution values
plot_pollution_map <- function(data, period) {
  # Aggregate data by period
  aggregated_data <- aggregate(. ~ Longitude + Latitude, data = data, FUN = mean)
  
  # Create the map plot
  ggplot() +
    geom_polygon(data = map_limits, aes(x = long, y = lat, group = group), fill = "lightgray", color = "black") +
    geom_point(data = aggregated_data, aes(x = Longitude, y = Latitude, color = PM1), size = 3) +
    scale_color_gradient(low = "green", high = "red", name = "PM1") +
    labs(title = paste("Average PM1", period)) +
    theme_minimal()
}

# Plot the maps for different periods
plot_all_time <- plot_pollution_map(pollution_data, "All Time")
plot_monthly <- plot_pollution_map(pollution_data, "Monthly")
plot_daily <- plot_pollution_map(pollution_data, "Daily")
plot_hourly <- plot_pollution_map(pollution_data, "Hourly")
```

### Plots using OpenAir

```{r}
# TimePlots
class(date)
timePlot(modPM_met, pollutant = c("pm1", "pm25", "pm10"))
timePlot(modPM_met, pollutant = c("wd", "ws"))
```

```{r}
# Diurnals
timeVariation(modPM_met, pollutant = c("pm1", "pm25", "pm10"), local.tz= "America/New_York") 

```

```{r}
# Remove zero wind speeds
modPM_met_nz <- modPM_met
modPM_met_nz$ws[modPM_met_nz$ws==0]<-NA

```

```{r}
# Polar Plots - IQR
  polarPlot(modPM_met_nz, pollutant = "pm1", limits = c(0,15), main = paste0("Roxbury ", "PM1 Polar Plot", sep= " "))
  polarPlot(modPM_met_nz, pollutant = "pm25", limits = c(0,20), main = paste0("Roxbury ", "PM2.5 Polar Plot", sep= " "))
  polarPlot(modPM_met_nz, pollutant = "pm10", limits = c(0,30),  main = paste0("Roxbury ", "PM10 Polar Plot", sep= " "))
```

```{r}
# Polar Plots as weighted mean
  polarPlot(modPM_met_nz, pollutant = "pm1", statistic = "weighted.mean", normalise = TRUE,main = paste0("Roxbury ", "PM1 Polar Plot", sep= " "))
  polarPlot(modPM_met_nz, pollutant = "pm25",statistic = "weighted.mean", normalise = TRUE, main = paste0("Roxbury ", "PM2.5 Polar Plot", sep= " "))
  polarPlot(modPM_met_nz, pollutant = "pm10",statistic = "weighted.mean", normalise = TRUE, main = paste0("Roxbury ", "PM10 Polar Plot", sep= " "))
```

```{r}
  polarFreq(modPM_met_nz, pollutant = "pm1",  statistic = "weighted.mean", main = paste0("Roxbury ", "PM1 Polar Plot", sep= " "))
  polarFreq(modPM_met_nz, pollutant = "pm25", statistic = "weighted.mean",  main = paste0("Roxbury ", "PM2.5 Polar Plot", sep= " "))
  polarFreq(modPM_met_nz, pollutant = "pm10", statistic = "weighted.mean",  main = paste0("Roxbury ", "PM10 Polar Plot", sep= " "))
```

```{r}
polarFreq(modPM_met_nz, pollutant = "pm1", ws.int = 30, statistic = "weighted.mean", offset = 80, trans = FALSE, col = "heat", main = paste0("Roxbury ", "PM1 Polar Plot", sep= " "))
polarFreq(modPM_met_nz, pollutant = "pm25", ws.int = 30, statistic = "weighted.mean", offset = 80, trans = FALSE, col = "heat", main = paste0("Roxbury ", "PM2.5 Polar Plot", sep= " "))
polarFreq(modPM_met_nz, pollutant = "pm10", ws.int = 30, statistic = "weighted.mean", offset = 80, trans = FALSE, col = "heat", main = paste0("Roxbury ", "PM10 Polar Plot", sep= " "))
```

```{r}
polarFreq(modPM_met_nz, pollutant = "pm1", ws.int = 30, statistic = "frequency", offset = 80, trans = FALSE, col = "heat")
polarFreq(modPM_met_nz, pollutant = "pm25", ws.int = 30, statistic = "frequency", offset = 80, trans = FALSE, col = "heat")
polarFreq(modPM_met_nz, pollutant = "pm10", ws.int = 30, statistic = "frequency", offset = 80, trans = FALSE, col = "heat")
```

```{r}
polarFreq(modPM_met_nz, pollutant = "pm1", ws.int = 30,  offset = 80, trans = FALSE, col = "heat")
polarFreq(modPM_met_nz, pollutant = "pm25", ws.int = 30,  offset = 80, trans = FALSE, col = "heat")
polarFreq(modPM_met_nz, pollutant = "pm10", ws.int = 30,  offset = 80, trans = FALSE, col = "heat")
```

```{r}
modPM_met_nz$lat <- 42.440740
modPM_met_nz$long <- -70.969658

polarMap(modPM_met_nz,pollutant = "pm1",statistic = "weighted.mean",x = "ws",latitude = "lat",longitude = "long", provider    = "OpenStreetMap",cols = "jet",alpha = 1,key = FALSE,iconWidth = 200,iconHeight = 200,fig.width = 4,fig.height = 4)

polarMap(modPM_met_nz,pollutant = "pm25",statistic = "weighted.mean",x = "ws",latitude = "lat",longitude = "long", provider    = "OpenStreetMap",cols = "jet",alpha = 1,key = FALSE,iconWidth = 200,iconHeight = 200,fig.width = 4,fig.height = 4)

polarMap(modPM_met_nz,pollutant = "pm10",statistic = "weighted.mean",x = "ws",latitude = "lat",longitude = "long", provider    = "OpenStreetMap",cols = "jet",alpha = 1,key = FALSE,iconWidth = 200,iconHeight = 200,fig.width = 4,fig.height = 4)
```

```{r}
# Polar Clusters
polarCluster(modPM_met, pollutant = "pm1", n.clusters = 2:10, cols= "Set2", main = paste0("Roxbury"))

polarCluster(modPM_met, pollutant = "pm25", n.clusters = 2:10, cols= "Set2", main = paste0("Roxbury"))

polarCluster(modPM_met, pollutant = "pm10", n.clusters = 2:10, cols= "Set2", main = paste0("Roxbury"))
```

```{r}
# Polar Annulus
polarAnnulus(modPM_met_nz, poll = "pm10", period = "trend", main = "Trend")
polarAnnulus(modPM_met_nz, poll = "pm10", period = "season", main = "Season")
polarAnnulus(modPM_met_nz, poll = "pm10", period = "weekday", main = "Weekday")
polarAnnulus(modPM_met_nz, poll = "pm10",period = "hour", main = "Hour")

```
