---
title: "PM Stacks"
author: "Cherry Pham"
date: "2023-07-20"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

This Markdown file plots the average pollutant levels by sensor per
month and the total pollution levels for the year.

The resulting figures are saved in `ambient-summer23/figures/pm_stacks/`
and can also be visualized below. Simply press `Ctrl + Alt + R` to run
all and get the figures. I recommend looking at the functions in this file -
`./helpers/ploting_helpers.R` to explore your options and modify stuff.

## INITAL SETTINGS

```{r setup, include=FALSE}
# Check and install required packages if necessary
packages <- c("lubridate", "dplyr", "tidyverse", "wesanderson", "svglite")
install.packages(packages[!sapply(packages, requireNamespace, quietly = TRUE)])

# Load required packages for data manipulation and analysis
invisible(sapply(packages, library, character.only = TRUE))

# Import helper functions
source("./helpers/plotting_helpers.R")

# Set options
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

## DATA LOADING AND CLEANING

```{r}
# Load data frame
load("../data/graphableData.RData")
```

## PLOTTING POLLUTION PER SENSOR

Create bar graphs with each bar representing the pollutants from a
single sensor

```{r}
# Create a subset of the data for PM1, PM2.5, and PM10
pm_subset <- mod_met[, c("sitename", "pm1", "pm25", "pm10", "date")]

# Convert the date column to a month format
pm_subset$date <- as.Date(pm_subset$date)
pm_subset$month <- format(pm_subset$date, "%Y-%m")

# Reshape the data into long format
pm_subset_long <-
  tidyr::pivot_longer(
    pm_subset,
    cols = c(pm1, pm25, pm10),
    names_to = "pollutant",
    values_to = "value"
  )

# Specify the order of pollutants in the stack
pm_subset_long$pollutant <-
  factor(pm_subset_long$pollutant, levels = c("pm1", "pm25", "pm10"))

# Calculate monthly averages
monthly_averages <- pm_subset_long %>%
  group_by(sitename, month, pollutant) %>%
  summarise(value = mean(value))

# Assign colors to each pollutant
color_palette <- wes_palette(n = 3, name = "GrandBudapest1", type = "continuous")
color_mapping <- c("pm10" = color_palette[1],
                   "pm25" = color_palette[3],
                   "pm1" = color_palette[2])


```

```{r}
# Find the sensor with the maximum sum of PM1, PM2.5, and PM10 values
  # for each month, then find the overall maximum value for y-axis scaling
max_sensors <- find_max_pollutant(monthly_averages)
y_axis_max <- max(max_sensors$sum_value)

# Generate montly graphs
monthly_graphs <- generate_monthly_graphs(monthly_averages, y_axis_max, color_mapping)

# Print the monthly graphs
for (graph in monthly_graphs) {
  print(graph)
}
```

```{r}
# Specify the year you want to filter
specified_year <- 2022

# Sum PM1, PM2.5, and PM10 per sensor by year
yearly_sums <- pm_subset_long %>%
  filter(lubridate::year(date) == specified_year) %>%
  group_by(sitename, year = lubridate::year(date), pollutant) %>%
  summarise(value = sum(value))

```

```{r}
yearly_graph <- generate_yearly_graph(yearly_sums, color_mapping)
print(yearly_graph) # There's a full window button that helps the preview look less nasty
```

```{r}
# # Save the monthly graphs as jpg files
# for (i in seq_along(monthly_graphs)) {
#   filename <- paste0("../figures/pm_stacks/monthly_graph_", i, ".jpg")
#   save_plot_as_jpg(monthly_graphs[[i]], filename)
# }
# # Save the yearly graph as a jpg file
# save_plot_as_jpg(total_plot, "../figures/pm_stacks/total_pollutant_levels.jpg")
```
