---
title: "Polar Plots"
date: "2023-07-17"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

This file generates PM1, PM2.5, and PM10 polar plots on a map showing to
each sensor's location and the corresponding pollution readings.

Simply press `Ctrl + Alt + R` to run all and get the figures. Go to `./helpers/`
and run the file `exporting_html.R` according to instructions in the file to get
the html export of this file. Note that this file is quite memory taxing so make
sure that your R memory (the pie chart at the top of the Environment tab) is
green to make sure nothing crashes. 

## INITIAL SETTINGS

```{r setup, include=FALSE}
# Check and install required packages if necessary
packages <- c("openair", "openairmaps", "leaflet", "dplyr")
install.packages(packages[!sapply(packages, requireNamespace, quietly = TRUE)])

# Load required packages for data manipulation and analysis
invisible(sapply(packages, library, character.only = TRUE))

# Set options
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

## DATA LOADING AND CLEANING

```{r}
# Load data frame
load("../data/graphableData.RData")

# Filter the data to include only the desired time range (time range change)
start_date <- as.Date("2022-07-01")
end_date <- as.Date("2023-02-28")
mod_met_filtered <- mod_met[mod_met$date >= start_date & mod_met$date <= end_date, ]

# Make a smaller dataset so things run faster
mod_met_filtered <- mod_met_filtered[, c("date", "sn", "pm1", "pm25", "pm10", "ws", "wd", "lat", "lon")]
```

## POLAR PLOTTING

### Plotting Only PM1

```{r}
polarMap(
  mod_met_filtered,
  pollutant = c("pm1"),
  latitude = "lat",
  longitude = "lon",
  limits = c(0, 5),
  key = TRUE,
  provider = "OpenStreetMap")
```

### Plotting Only PM2.5

```{r}
polarMap(
  mod_met_filtered,
  pollutant = c("pm25"),
  latitude = "lat",
  longitude = "lon",
  limits = c(0, 12),
  key = TRUE,
  provider = "OpenStreetMap")
```

### Plotting Only PM10

```{r}
polarMap(
  mod_met_filtered,
  pollutant = c("pm10"),
  latitude = "lat",
  longitude = "lon",
  limits = c(5, 35),
  key = TRUE,
  provider = "OpenStreetMap")
```

### Plotting all PMs Polar Plots

Create a map with the average polar plots w/ of each pollution values
(PM1, PM2.5, PM10) - toggle/normalize.

```{r}
polarMap(
  mod_met_filtered,
  pollutant = c("pm1", "pm25", "pm10"),
  latitude = "lat",
  longitude = "lon",
  normalize = TRUE,
  key = TRUE,
  provider = "OpenStreetMap")
```
