---
title: "Sanity Checks"
author: "Cherry Pham"
date: "2023-07-06"
output: html_document
---

```{r}
# Check and install required packages if necessary
packages <- c(
  "lubridate",   # Package for working with dates and times
  "data.table",  # Package for efficient data manipulation and processing
  "dplyr",       # Package for data manipulation and transformation
  "openair",     # Package for analyzing air pollution data
  # "openairmaps", # Package for accessing open air pollution data
  "stringr",     # Package for string manipulation
  "baseline",    # Package for baseline modeling and adjustment
  "purrr",       # Package for functional programming
  "tidyverse",   # Meta-package that includes several tidyverse packages for data science
  "openairmaps", # Package for accessing open air pollution data
  "lattice",
  "rstudioapi"
)
install.packages(packages[!sapply(packages, requireNamespace, quietly = TRUE)])

# Load required packages for data manipulation and analysis
invisible(sapply(packages, library, character.only = TRUE))

# Set options
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

## VISUALIZING RESULTS

```{r}
# List of devices
# device_ids <- c(
#   "MOD-00024",
#   "MOD-00025",
#   "MOD-00026",
#   "MOD-00027",
#   "MOD-00028",
#   "MOD-PM-00141",
#   "MOD-PM-00211",
#   "MOD-PM-00212",
#   "MOD-PM-00213",
#   "MOD-PM-00214",
#   "MOD-PM-00216",
#   "MOD-PM-00217",
#   "MOD-PM-00221",
#   "MOD-PM-00222",
#   "MOD-PM-00224",
#   "MOD-PM-00226",
#   "MOD-PM-00230",
#   "MOD-PM-00231"
# )

# View one device
view(read.csv("./data/Quant-AQ/MOD-PM/MOD-00024.csv"))

# View multiple devices
# for (device_id in device_ids) {
#   view(read.csv(paste0("./data/Quant-AQ/MOD-PM/", device_id,".csv")))
# }

```

## GRAPH THEMES

### ggplot2:

```{r}
library(ggplot2)
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()
```

### Custom ggplot2:

```{r}
library(ggplot2)
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  theme_bw()
```

### Custom lattice:

```{r}
library(lattice)
trellis.par.set(
  strip.background = list(col = "lightgray"),
  axis.line = list(col = "black"),
  par.strip.text = list(col = "black")
)

xyplot(mpg ~ wt, data = mtcars, type = c("p", "smooth"))

```

### Base R graphics:

```{r}
plot(mpg ~ wt, data = mtcars,
     col = "blue", pch = 16,
     xlab = "Weight", ylab = "MPG",
     main = "Scatterplot of MPG vs. Weight")
```

### Other options

1.  ggplot2 Themes and Customization:

-   ggplot2: In addition to using pre-defined themes, you can further customize your ggplot2 graphs by modifying various aspects such as axes, legends, titles, and annotations. You can use functions like \`theme()\`, \`labs()\`, and \`annotate()\` to make specific changes. Additionally, you can modify the appearance of individual plot elements using the \`element\_\*()\` functions. For example, \`element_line()\` allows you to modify line properties, \`element_text()\` allows you to change text properties, and so on.

2.  Plotly and Interactive Graphs:

-   plotly: The plotly package allows you to create interactive graphs in R. You can use it to create dynamic and visually appealing visualizations with features like zooming, panning, tooltips, and more. Plotly supports a wide range of graph types, including scatter plots, bar charts, heatmaps, and 3D plots. It also provides options for customization, such as modifying colors, adding hover effects, and specifying layout configurations.

3.  Complex Visualization Libraries:

-   Highcharter: The Highcharter package provides an interface to Highcharts, a powerful JavaScript library for interactive charts. It allows you to create advanced charts like treemaps, network graphs, and heatmaps.

-   NetworkD3: This package enables the creation of interactive network visualizations, including force-directed layouts, hierarchical edge bundling, and Sankey diagrams.

-   ggvis: ggvis is an interactive visualization package built on top of ggplot2. It allows you to create interactive graphs using reactive programming concepts, similar to Shiny applications.

4.  Custom Graphics with base R:

-   For advanced customization in base R graphics, you can explore functions like \`par()\` and \`plot.default()\` to control various graphical parameters. You can modify settings such as colors, line types, axis labels, plot margins, and more.

### Mapping w. R

```{r}
polarMap(
  openairmaps::polar_data,
  pollutant = c("nox", "no2"),
  latitude = "lat",
  longitude = "lon",
  provider = "CartoDB.Positron")
```

### Plots using OpenAir

```{r}
# TimePlots
class(date)
timePlot(modPM_met, pollutant = c("pm1", "pm25", "pm10"))
timePlot(modPM_met, pollutant = c("wd", "ws"))
gc()
```

```{r}
# Diurnals
timeVariation(modPM_met, pollutant = c("pm1", "pm25", "pm10"), local.tz= "America/New_York") 
gc()
```

```{r}
# Remove zero wind speeds
modPM_met_nz <- modPM_met
modPM_met_nz$ws[modPM_met_nz$ws==0]<-NA
gc()
```

```{r}
# Polar Plots - IQR
  polarPlot(modPM_met_nz, pollutant = "pm1", limits = c(0,15), main = paste0("Roxbury ", "PM1 Polar Plot", sep= " "))
  polarPlot(modPM_met_nz, pollutant = "pm25", limits = c(0,20), main = paste0("Roxbury ", "PM2.5 Polar Plot", sep= " "))
  polarPlot(modPM_met_nz, pollutant = "pm10", limits = c(0,30),  main = paste0("Roxbury ", "PM10 Polar Plot", sep= " "))
  gc()
```

```{r}
# Polar Plots as weighted mean
  polarPlot(modPM_met_nz, pollutant = "pm1", statistic = "weighted.mean", normalise = TRUE,main = paste0("Roxbury ", "PM1 Polar Plot", sep= " "))
  polarPlot(modPM_met_nz, pollutant = "pm25",statistic = "weighted.mean", normalise = TRUE, main = paste0("Roxbury ", "PM2.5 Polar Plot", sep= " "))
  polarPlot(modPM_met_nz, pollutant = "pm10",statistic = "weighted.mean", normalise = TRUE, main = paste0("Roxbury ", "PM10 Polar Plot", sep= " "))
  gc()
```

```{r}
  polarFreq(modPM_met_nz, pollutant = "pm1",  statistic = "weighted.mean", main = paste0("Roxbury ", "PM1 Polar Plot", sep= " "))
  polarFreq(modPM_met_nz, pollutant = "pm25", statistic = "weighted.mean",  main = paste0("Roxbury ", "PM2.5 Polar Plot", sep= " "))
  polarFreq(modPM_met_nz, pollutant = "pm10", statistic = "weighted.mean",  main = paste0("Roxbury ", "PM10 Polar Plot", sep= " "))
  gc()
```

```{r}
polarFreq(modPM_met_nz, pollutant = "pm1", ws.int = 30, statistic = "weighted.mean", offset = 80, trans = FALSE, col = "heat", main = paste0("Roxbury ", "PM1 Polar Plot", sep= " "))
polarFreq(modPM_met_nz, pollutant = "pm25", ws.int = 30, statistic = "weighted.mean", offset = 80, trans = FALSE, col = "heat", main = paste0("Roxbury ", "PM2.5 Polar Plot", sep= " "))
polarFreq(modPM_met_nz, pollutant = "pm10", ws.int = 30, statistic = "weighted.mean", offset = 80, trans = FALSE, col = "heat", main = paste0("Roxbury ", "PM10 Polar Plot", sep= " "))
gc()
```

```{r}
polarFreq(modPM_met_nz, pollutant = "pm1", ws.int = 30, statistic = "frequency", offset = 80, trans = FALSE, col = "heat")
polarFreq(modPM_met_nz, pollutant = "pm25", ws.int = 30, statistic = "frequency", offset = 80, trans = FALSE, col = "heat")
polarFreq(modPM_met_nz, pollutant = "pm10", ws.int = 30, statistic = "frequency", offset = 80, trans = FALSE, col = "heat")
gc()
```

```{r}
polarFreq(modPM_met_nz, pollutant = "pm1", ws.int = 30,  offset = 80, trans = FALSE, col = "heat")
polarFreq(modPM_met_nz, pollutant = "pm25", ws.int = 30,  offset = 80, trans = FALSE, col = "heat")
polarFreq(modPM_met_nz, pollutant = "pm10", ws.int = 30,  offset = 80, trans = FALSE, col = "heat")
gc()
```

```{r}
modPM_met_nz$lat <- 42.440740
modPM_met_nz$long <- -70.969658

polarMap(modPM_met_nz,pollutant = "pm1",statistic = "weighted.mean",x = "ws",latitude = "lat",longitude = "long", provider    = "OpenStreetMap",cols = "jet",alpha = 1,key = FALSE,iconWidth = 200,iconHeight = 200,fig.width = 4,fig.height = 4)

polarMap(modPM_met_nz,pollutant = "pm25",statistic = "weighted.mean",x = "ws",latitude = "lat",longitude = "long", provider    = "OpenStreetMap",cols = "jet",alpha = 1,key = FALSE,iconWidth = 200,iconHeight = 200,fig.width = 4,fig.height = 4)

polarMap(modPM_met_nz,pollutant = "pm10",statistic = "weighted.mean",x = "ws",latitude = "lat",longitude = "long", provider    = "OpenStreetMap",cols = "jet",alpha = 1,key = FALSE,iconWidth = 200,iconHeight = 200,fig.width = 4,fig.height = 4)
gc()
```

```{r}
# Polar Clusters
polarCluster(modPM_met, pollutant = "pm1", n.clusters = 2:10, cols= "Set2", main = paste0("Roxbury"))
polarCluster(modPM_met, pollutant = "pm25", n.clusters = 2:10, cols= "Set2", main = paste0("Roxbury"))
polarCluster(modPM_met, pollutant = "pm10", n.clusters = 2:10, cols= "Set2", main = paste0("Roxbury"))
gc()
```

```{r}
# Polar Annulus
polarAnnulus(modPM_met_nz, poll = "pm10", period = "trend", main = "Trend")
polarAnnulus(modPM_met_nz, poll = "pm10", period = "season", main = "Season")
polarAnnulus(modPM_met_nz, poll = "pm10", period = "weekday", main = "Weekday")
polarAnnulus(modPM_met_nz, poll = "pm10",period = "hour", main = "Hour")
gc()
```
