---
title: "Toggle Map"
author: "Cherry Pham"
date: "2023-07-22"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

This file creates maps using the leaflet function and add data in
representative circles.

Simply press `Ctrl + Alt + R` to run all and get the figures. Go to
`ambient-summer23/helpers/` and run the file `exporting_html.R`
according to instructions in the file to get the html export of this
file.

## INITIAL SETTINGS

```{r setup, include=FALSE}
# Check and install required packages if necessary
packages <- c("openair", "openairmaps", "leaflet", "dplyr")
install.packages(packages[!sapply(packages, requireNamespace, quietly = TRUE)])

# Load required packages for data manipulation and analysis
invisible(sapply(packages, library, character.only = TRUE))

# Set options
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

## DATA LOADING AND CLEANING

```{r}
# Load data frame
load("../data/graphableData.RData")
```

```{r}
print(mod_met)
```

```{r}
# Calculate the average of each PM
averagePM1 <- mean(mod_met$pm1)
averagePM25 <- mean(mod_met$pm25)
averagePM10 <- mean(mod_met$pm10)

# Set the threshold for significant difference
threshold <- 2

# Filter out rows where the column value is significantly bigger than the average
mod_met_filtered <- mod_met[mod_met$pm1 <= averagePM1 * threshold, ]
mod_met_filtered <- mod_met[mod_met$pm25 <= averagePM25 * threshold, ]
mod_met_filtered <- mod_met[mod_met$pm10 <= averagePM10 * threshold, ]

# Helper function
subset_by_month <- function(dataset, year, month) {
  # Check if the 'date' column exists in the dataset
  if ("date" %in% names(dataset)) {
    # Convert 'date' column to proper Date format (if needed)
    if (!inherits(dataset$date, "Date")) {
      dataset$date <- as.Date(dataset$date)
    }

    # Create a variable for the target month's start and end dates
    start_date <- as.Date(paste(year, month, "01", sep = "-"))
    end_date <- as.Date(paste(year, month + 1, "01", sep = "-")) - 1

    # Filter data for the specified month
    month_subset <- dataset[dataset$date >= start_date & dataset$date <= end_date, ]
    return(month_subset)
  }
}

# Filter the data to include only the desired time range
mod_met_subset <- subset_by_month(mod_met, year = 2022, month = 8)

# Calculate the average data for pm1, pm25, and pm10 per sensor
average_data <- mod_met_subset %>%
  group_by(sn) %>%
  summarize(
    avg_pm1 = mean(pm1),
    avg_pm25 = mean(pm25),
    avg_pm10 = mean(pm10),
    lon = mean(lon),
    lat = mean(lat)   
  ) 

# Also helper
fill_null_with_zero <- function(df) {
  df[is.na(df)] <- 0
  return(df)
}
average_data <- fill_null_with_zero(average_data)
```

## PLOTTING

### Circles indicating pollution on map

Create a map of for the average of each pollution values (PM1, PM2.5,
PM10) over certain periods of time (all time, monthly, daily, hourly)

```{r}
# library(wesanderson)
# 
# color_palette <-
#   wes_palette(n = 3, name = "GrandBudapest1", type = "continuous")
# 
# # Create the leaflet map
# map <- leaflet(average_data) %>%
#   addTiles()
# 
# # Display the map with all three layers using addLayersControl
# map_combined <- map %>%
#   addLayersControl(
#     overlayGroups = c("PM1", "PM2.5", "PM10"),
#     options = layersControlOptions(collapsed = FALSE, position = "topright")
#   ) %>%
#   addCircles(
#     lng = ~ lon,
#     lat = ~ lat,
#     fillOpacity = 0.5,
#     fillColor = color_palette[1],
#     weight = 0,
#     radius = ~ sqrt(avg_pm10) * 30,
#     popup = ~ paste("Sensor:", sn, "<br>PM10:", round(avg_pm10, 2)),
#     group = "PM10"
#   ) %>%
#   addCircles(
#     lng = ~ lon,
#     lat = ~ lat,
#     fillOpacity = 0.7,
#     fillColor = color_palette[3],
#     weight = 0,
#     radius = ~ sqrt(avg_pm25) * 30,
#     popup = ~ paste("Sensor:", sn, "<br>PM2.5:", round(avg_pm25, 2)),
#     group = "PM2.5"
#   ) %>%
#   addCircles(
#     lng = ~ lon,
#     lat = ~ lat,
#     fillOpacity = 0.9,
#     fillColor = color_palette[2],
#     weight = 0,
#     radius = ~ sqrt(avg_pm1) * 30,
#     popup = ~ paste("Sensor:", sn, "<br>PM1:", round(avg_pm1, 2)),
#     group = "PM1"
#   )
# 
# print(map_combined)
```

### Circles indicating heat on map

Create a map of for the average of each temperature values over certain
periods of time (monthly, weekly, daily)

```{r}
# Filter the data to include only the desired time range
mod_met_subset <- subset_by_month(mod_met, year = 2022, month = 8)

# Calculate the average data for pm1, pm25, and pm10 per sensor
mod_met_temp <- mod_met_subset %>%
  group_by(sn) %>%
  summarize(
    avg_temp = mean(tmpc),
    lon = mean(lon),
    lat = mean(lat)   
  ) 

# Create the leaflet map
map <- leaflet(mod_met_temp) %>%
  addTiles()

# Display the map with all three layers using addLayersControl
map_combined <- map %>%
  addCircles(
    lng = ~ lon,
    lat = ~ lat,
    fillOpacity = 0.7,
    fillColor = "orange",
    weight = 0,
    radius = ~ sqrt(avg_temp) * 30
  )
print(map_combined)
```
