---
title: "API Call"
author: "Cherry Pham"
date: "2023-06-20"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

This Markdown file is created with the purpose of obtaining sensor data
from Quant-AQ through API calls.

There is no prerequisites for using this file. If you are just looking
to quickly retrieve data, you can simply run this file through the
command `Ctrl + Alt + R`. The result of running this file is a data
folder containing data of MOD-PM sensors up to 07/31/2023. Press
`Ctrl + F` and find `(time range change)` to modify the time range to
retrieve data. If you wish to modify the data to be retrieved, please
read the data guide below.

## DATA GUIDE

A. **PACKAGES AND INITIAL SETTINGS**

B. **DEFINE VARIABLES**

-   You can find sensors information, data-retrieving time range, API
    key and API URL here.

C. **RETTRIEVING DATA INTO FOLDERS**
1.  Check if the appropriate data folders exist and create them if not.

2.  `process_device_id` function:

    a.  Loop over the date range and makes API calls for each date.

    b.  The API response is checked for errors and processed
        accordingly.

    c.  After processing all the dates for a sensor, merge the raw and
        final data based on the "timestamp_local" column into the data
        frame `merged_data_per_day`.

    d.  Flatten merged data by unnesting any nested data frames within
        it and append all data across time range for each sensor
        vertically into the data frame `merged_data_per_sensor`.

    e.  Attempt to write the `merged_data_per_sensor` data frame into a
        CSV file. If an error occurs during the writing process, append
        an error message to the "error_log.txt" file.

3.  Loop over each sensor deployment in the list `device_ids` to apply
    `process_device_id` to every device.

## PACKAGES AND INITIAL SETTINGS

```{r}
# Check and install required packages if necessary
packages <- c("httr", "jsonlite", "purrr", "dplyr", "tidyr")
install.packages(packages[!sapply(packages, requireNamespace, quietly = TRUE)])

# Load required packages
invisible(sapply(packages, library, character.only = TRUE))

# Explicit package function calls
httr::GET
httr::content
jsonlite::fromJSON
purrr::map_df
dplyr::as_data_frame
dplyr::filter
dplyr::lag

# Set knit options and working directory
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

## DEFINE VARIABLES

```{r}
base_url <- "https://api.quant-aq.com/device-api/v1/devices/"
api_key <- "VUOLNUDTM70QZ7Z9G0Z78XA7" # Cherry's API Key

# Define a time range to loop over (time range change)
time_range <-
  seq(as.Date("2022-07-01"), as.Date("2023-08-01"), by = "day")
time_range <- format(time_range, "%Y-%m-%d")

# Define a list of device IDs
device_ids <- c(
  "MOD-00024",
  "MOD-00025",
  "MOD-00026",
  "MOD-00027",
  "MOD-00028",
  "MOD-PM-00141",
  "MOD-PM-00211",
  "MOD-PM-00212",
  "MOD-PM-00213",
  "MOD-PM-00214",
  "MOD-PM-00216",
  "MOD-PM-00217",
  "MOD-PM-00221",
  "MOD-PM-00222",
  "MOD-PM-00224",
  "MOD-PM-00226",
  "MOD-PM-00230",
  "MOD-PM-00231"
)
```

## RETTRIEVING DATA INTO FOLDERS

```{r}
# Import helper function(s)
# Make sure that process_device_id.R is in this folder
source("./helpers/API_call_helpers.R")

# Create folder if not initiated
data_folder_path <- "../data/QuantAQ/MODPM/"

# Check if the folder exists and create the folder if it doesn't exist
if (!file.exists(data_folder_path)) {
  dir.create(data_folder_path, recursive = TRUE)
}

# Process each device ID over time range
for (device_id in device_ids) {
  process_device_id(device_id)
}
```
