---
title: "API Call"
author: "Cherry Pham"
date: "2023-06-20"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

This Markdown file is created with the purpose of obtaining sensor data
from Quant-AQ through API calls.

There is no prerequisites for using this file. If you are just looking
to quickly retrieve data, you can simply run this file through the
command `Ctrl + Alt + R`. The result of running this file is a data
folder containing data of MOD-PM sensors up to 07/31/2023. Press
`Ctrl + F` and find `(time range change)` to modify the time range to
retrieve data. If you wish to modify the data to be retrieved, please
read the data guide below.

## DATA GUIDE

This section describe the Outline (the links you can go to from the tab on the
right that's under the run button). Hopefully this makes the different sections'
content more easy to access. 

A. **PACKAGES AND INITIAL SETTINGS**

-   You can find packages installation and loading, helper function import and
    knit options here.

B. **DEFINE VARIABLES**

-   You can find sensors information, data-retrieving time range, API
    key and API URL here.

C. **RETRIEVING DATA INTO FOLDERS**

-   You can find the function that make the API call, create and name folders,
    and create and name files here.

## PACKAGES AND INITIAL SETTINGS

```{r}
# Check and install required packages if necessary
packages <- c("httr", "jsonlite", "purrr", "dplyr", "tidyr")
install.packages(packages[!sapply(packages, requireNamespace, quietly = TRUE)])

# Load required packages
invisible(sapply(packages, library, character.only = TRUE))

# Explicit package function calls
httr::GET
httr::content
jsonlite::fromJSON
purrr::map_df
dplyr::as_data_frame
dplyr::filter
dplyr::lag

# Import helper functions
source("./helpers/API_call_helpers.R")

# Set knit options
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

## DEFINE VARIABLES

```{r}
base_url <- "https://api.quant-aq.com/device-api/v1/devices/"
api_key <- "VUOLNUDTM70QZ7Z9G0Z78XA7" # Cherry's API Key

# Define a time range to loop over (time range change)
time_range <-
  seq(as.Date("2022-07-01"), as.Date("2023-08-01"), by = "day")
time_range <- format(time_range, "%Y-%m-%d")

# Define a list of device IDs
device_ids <- c(
  "MOD-00024",
  "MOD-00025",
  "MOD-00026",
  "MOD-00027",
  "MOD-00028",
  "MOD-PM-00141",
  "MOD-PM-00211",
  "MOD-PM-00212",
  "MOD-PM-00213",
  "MOD-PM-00214",
  "MOD-PM-00216",
  "MOD-PM-00217",
  "MOD-PM-00221",
  "MOD-PM-00222",
  "MOD-PM-00224",
  "MOD-PM-00226",
  "MOD-PM-00230",
  "MOD-PM-00231"
)
```

## RETTRIEVING DATA INTO FOLDERS

```{r}
# Create folder if not initiated
data_folder_path <- "../data/QuantAQ/MODPM/"

# Check if the folder exists and create the folder if it doesn't exist
if (!file.exists(data_folder_path)) {
  dir.create(data_folder_path, recursive = TRUE)
}

# Process each device ID over time range
for (device_id in device_ids) {
  process_device_id(device_id)
}
```
