---
title: "Data Breakdown"
author: "Cherry Pham"
date: "2023-07-10"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

This Markdown file is created to break existing .csv files into monthly
files.

Currently, this file is made to breakdown **Meteorology data**
downloaded from the Iowa Environmental Mesonet and **Quant-AQ data**
retrieved using `API_call.Rmd`. See the **FETCHING DATA TO IMPORT**
session in this file for more details. As a side note, you want to, you
can also rewrite `API_call.Rmd` to not merge monthly files. Press
`Ctrl + Alt + R` to run all and breakdown all meteorology data and
Quant-AQ data. Make sure that you have all the needed files so that the
code doesn't error out.

## PACKAGES AND INITAL SETTINGS

```{r}
# Check and install required packages if necessary
packages <- c("lubridate", "data.table")
install.packages(packages[!sapply(packages, requireNamespace, quietly = TRUE)])

# Load required packages for data manipulation and analysis
invisible(sapply(packages, library, character.only = TRUE))

# Import helper function
source("./helpers/data_breakdown_helpers.R")

# Set knit options
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

## FETCHING DATA TO IMPORT

In order for this markdown file to function, you need all your Quant-AQ
and Meteorology data downloaded as csv. To skip this session (or this
entire data acquisition step), go to the data folder branch of [this
repository](https://github.com/airpartners/ambient-summer23/tree/datafolder)
to download a pre-made data folder. If you want to modify data/download
data for yourself, refer to the below instructions within this session.

### Quant-AQ Data

Refer to the Markdown file `API_call.Rmd` in this folder.

### Meteorology Data

The meteorology data used in some analysis below are taken from the
[Iowa Environmental
Mesonet](https://mesonet.agron.iastate.edu/request/download.phtml?network=MA_ASOS).
Note that the reference meteorology data is taken from a station in East
Boston, which is relatively correct for Roxbury.

Below are the downloading instructions taken from the [EB Gibson Park
Analysis](https://github.com/scott-hersey/EB_AQ_Network/blob/master/Gibson%20Park%20Analysis.Rmd)
with some modification.

Select the network: \* Massachusett ASOS

Select the following station: \* [BOS] BOSTON/LOGAN INTL

Select the following variables:

-   Air Temperature (C)

-   Wind Direction

-   Wind Speed [mph]

Select the date range: `7/1/2022 to 8/1/2022` (this dataset is not
inclusive of the last date)

Select this timezone: `America/New_York`

Use the following download options: `comma delimited` ,

`no latitude/longitude vectors` ,

`no elevation` ,

`represent missing data with blank string` ,

`denote trace with blank string`

Create a folder for the met data: ~/data/Metdata/raw, and add the downloaded data file there.

## IMPORT DATA AND PRE-PROCESSING

Change the time range if you only want to break down data from a
specific time range (to optimize speed). `Ctrl + F` the following: (time
range change). Make sure that the helper file `data_breakdown_helpers.R`
is in the helpers folder and the cell in `PACKAGES AND INITIAL SETTINGS`
is ran.

### Quant-AQ Data

```{r}
# Define time range (time range change)
start_date <- ymd("2022-07-01")
end_date <- ymd("2023-08-01")

# Define a list of device IDs
device_ids <- c(
  "MOD-00024",
  # "MOD-00025",
  # "MOD-00026",
  # "MOD-00027",
  # "MOD-00028",
  # "MOD-PM-00141",
  # "MOD-PM-00211",
  # "MOD-PM-00212",
  # "MOD-PM-00213",
  # "MOD-PM-00214",
  # "MOD-PM-00216",
  # "MOD-PM-00217",
  # "MOD-PM-00221",
  # "MOD-PM-00222",
  # "MOD-PM-00224",
  # "MOD-PM-00226",
  "MOD-PM-00230",
  "MOD-PM-00231"
)

for (sensor_id in device_ids) {
  # Define file paths
  file_name <- paste0("../data/QuantAQ/MODPM/", sensor_id, ".csv")
  output_folder <-
    paste0("../data/QuantAQ/MODPM/brokenDown/", sensor_id)
  # Load data
  data <- load_csv_data(file_name, output_folder)
  # Convert timestamp_local column to datetime
  data$timestamp_local <- ymd_hms(data$timestamp_local)
  name <- sensor_id
  # Breakdown csv
  breakdown_csv(data, name, start_date, end_date)
}
```

### Meteorology Data

```{r}
# Define time range (time range change)
start_date <- ymd("2022-07-01")
end_date <- ymd("2023-08-01")

# Define file paths
file_name <- "../data/Metdata/raw/metdata.csv"
output_folder <- "../data/Metdata/brokenDown/"

# Load data
data <- load_csv_data(file_name, output_folder)
# Convert valid (time) column to datetime
data$valid <- parse_date_time(data$valid, orders = "ymd HMS", truncated = 6)
data$timestamp_local <- data$valid # Name change to fit with function
name <- "metdata"
# Breakdown csv
breakdown_csv(data, name, start_date, end_date)
```
