---
title: "Data Processing"
author: "Cherry Pham"
date: "2023-07-17"
output: html_document
---

[Markdown file docstrings]

## INITAL SETTINGS

```{r setup, include=FALSE}
# Check and install required packages if necessary
packages <- c(
  "lubridate",   # Package for working with dates and times
  "dplyr",       # Package for data manipulation and transformation
  "purrr",       # Package for functional programming
  "tidyverse"   # Meta-package for data science
)
install.packages(packages[!sapply(packages, requireNamespace, quietly = TRUE)])

# Load required packages for data manipulation and analysis
invisible(sapply(packages, library, character.only = TRUE))

# Set options
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

# Suppress Warnings (This is a me thing, comment this session out to see the conflicting functions)
suppressPackageStartupMessages(library(maps))
```

## FETCHING DATA TO IMPORT

To skip this session, go to the data folder branch of [this repository](https://github.com/airpartners/ambient-summer23/tree/datafolder) to download a pre-made data folder. If you want to modify data/download data for yourself, refer to the below instructions within this session.

### Quant-AQ Data

Refer to the Markdown file API_call.Rmd in this folder. The data should be in ./data/Quant-AQ/MOD-PM/ if nothing goes wrong.

### Meteorology Data

The meteorology data used in some analysis below are taken from the [Iowa Environmental Mesonet](https://mesonet.agron.iastate.edu/request/download.phtml?network=MA_ASOS).

Below are the downloading instructions taken from the [EB Gibson Park Analysis](https://github.com/scott-hersey/EB_AQ_Network/blob/master/Gibson%20Park%20Analysis.Rmd).

Select the following station: \* [BOS] BOSTON/LOGAN INTL

Select the following variables:

-   Temperature (C)

-   Wind Direction

-   Wind Speed [mph]

Select the date range: `7/1/2022 to 2/28/2022` (this dataset is not inclusive of the last date)

Select this timezone: `America/New_York`

Use the following download options: `comma delimited` ,

`no latitude/longitude vectors` ,

`no elevation` ,

`represent missing data with blank string` ,

`denote trace with blank string`

## IMPORT DATA AND PRE-PROCESSING

### Quant-AQ Data

```{r}
# Read in csv files from data folder
device_ids <- c(
  "MOD-00024",
  "MOD-00025",
  "MOD-00026",
  "MOD-00027",
  "MOD-00028",
  "MOD-PM-00141",
  "MOD-PM-00211",
  "MOD-PM-00212",
  "MOD-PM-00213",
  "MOD-PM-00214",
  "MOD-PM-00216",
  "MOD-PM-00217",
  "MOD-PM-00221",
  "MOD-PM-00222",
  "MOD-PM-00224",
  "MOD-PM-00226",
  "MOD-PM-00230",
  "MOD-PM-00231"
)

modPM <- suppressMessages({
  data <- device_ids %>%
    map_dfr(function(sensor_id) {
      paths <- sprintf(
        "./data/Quant-AQ/MOD-PM/broken-down/%s/%s-2022-%02d.csv",
        sensor_id, sensor_id, 7:12
      )
      map_dfr(paths, function(path) {
        suppressWarnings(
          tryCatch(
            read.csv(path),
            error = function(e) NULL
          )
        )
      })
    })
})
```

```{r}
# Rename to match manually downloaded datasets
modPM <- modPM %>%
  rename(
    timestamp = timestamp.x,
    sn = sn.x,
    bin0 = opc.bin0,
    bin1 = opc.bin1,
    bin2 = opc.bin2,
    bin3 = opc.bin3,
    bin4 = opc.bin4,
    bin5 = opc.bin5,
    bin6 = opc.bin6,
    bin7 = opc.bin7,
    bin8 = opc.bin8,
    bin9 = opc.bin9,
    bin10 = opc.bin10,
    bin11 = opc.bin11,
    bin12 = opc.bin12,
    bin13 = opc.bin13,
    bin14 = opc.bin14,
    bin15 = opc.bin15,
    bin16 = opc.bin16,
    bin17 = opc.bin17,
    bin18 = opc.bin18,
    bin19 = opc.bin19,
    bin20 = opc.bin20,
    bin21 = opc.bin21,
    bin22 = opc.bin22,
    bin23 = opc.bin23,
    opcn3_pm1 = opc.pm1,
    opcn3_pm25 = opc.pm25,
    opcn3_pm10 = opc.pm10,
    pm1_env = neph.pm1,
    pm25_env = neph.pm25,
    pm10_env = neph.pm10,
    neph_csat = neph.cscat,
    co_we = gases.co.we,
    co_ae = gases.co.ae,
    co_diff = gases.co.diff,
    no_we = gases.no.we,
    no_ae = gases.no.ae,
    no_diff = gases.no.diff,
    no2_we = gases.no2.we,
    no2_ae = gases.no2.ae,
    no2_diff = gases.no2.diff,
    o3_we = gases.o3.we,
    o3_ae = gases.o3.ae,
    ox_diff = gases.o3.diff,
    pm1_model_id = modelpmpm1,
    pm25_model_id = modelpmpm25,
    pm10_model_id = modelpmpm10
  )
```

```{r}
# Define the latitude and longitude values for each sensor
sn_lat <- c(
  "MOD-00024" = 42.33155829160215, "MOD-00025" = 42.32862119,
  "MOD-00026" = 42.325756, "MOD-00027" = 42.32012,
  "MOD-00028" = 42.3289, "MOD-PM-00141" = 42.3328,
  "MOD-PM-00211" = 42.3207, "MOD-PM-00212" = 42.326244,
  "MOD-PM-00213" = 42.3145978, "MOD-PM-00214" = 42.315932,
  "MOD-PM-00216" = 42.3294, "MOD-PM-00217" = 42.32355,
  "MOD-PM-00221" = 42.312182, "MOD-PM-00222" = 42.331,
  "MOD-PM-00224" = 42.32012, "MOD-PM-00226" = 42.3226,
  "MOD-PM-00230" = 42.3199, "MOD-PM-00231" = 42.3161
)

sn_lon <- c(
  "MOD-00024" = -71.08360367349657, "MOD-00025" = -71.086748391,
  "MOD-00026" = -71.068577, "MOD-00027" = -71.0679,
  "MOD-00028" = -71.0815, "MOD-PM-00141" = -71.097597,
  "MOD-PM-00211" = -71.0715, "MOD-PM-00212" = -71.081156,
  "MOD-PM-00213" = -71.0971196, "MOD-PM-00214" = -71.087439,
  "MOD-PM-00216" = -71.0777, "MOD-PM-00217" = -71.08866,
  "MOD-PM-00221" = -71.071364, "MOD-PM-00222" = -71.086,
  "MOD-PM-00224" = -71.0679, "MOD-PM-00226" = -71.0738,
  "MOD-PM-00230" = -71.0752, "MOD-PM-00231" = -71.0780
)

# Define the site names for each sensor
sn_siteName <- c(
  "MOD-00024" = "ACE Parking Lot", "MOD-00025" = "Hale Elem",
  "MOD-00026" = "The Base", "MOD-00027" = "Indigo Block",
  "MOD-00028" = "Hibernian Playgr.", "MOD-PM-00141" = "Tobin Comm Ctr",
  "MOD-PM-00211" = "Dudley Vil Playgr", "MOD-PM-00212" = "Dearborn Academy",
  "MOD-PM-00213" = "Rafael Hernandez", "MOD-PM-00214" = "Trotter Elem",
  "MOD-PM-00216" = "Orchard Gardens", "MOD-PM-00217" = "Residence Obierne Pl.",
  "MOD-PM-00221" = "Residence Baker Ave.", "MOD-PM-00222" = "Beryl Gardens",
  "MOD-PM-00224" = "Indigo Block", "MOD-PM-00226" = "Dudley Greenhouse",
  "MOD-PM-00230" = "Trina Persad Playgr", "MOD-PM-00231" = "Cardinal Medeiros"
)

# Add "lat", "lon", "siteName" columns to the existing data frame
modPM$lat <- sn_lat[modPM$sn]
modPM$lon <- sn_lon[modPM$sn]
modPM$siteName <- sn_siteName[modPM$sn]

print(modPM)
```

```{r}
# Combine pm1 bins
modPM$pm1num <- modPM$bin0 + modPM$bin1 + modPM$bin2

# Add a formatted date column
modPM$date <- as.POSIXct(strptime(modPM$timestamp_local, format = "%Y-%m-%d %H:%M:%S", tz = "America/New_York"))

# Delete unnecessary columns
columns_to_delete <- c(
  "url.x", "url.y", "sn.y", "timestamp.y", "geo.lat", "geo.lon", "met.rh", "met.temp", "met.wd", "met.ws",
  "geolat", "geolon", "metrh", "geo.xlat", "geo.xlon", "met.xrh", "met.xtemp", "nephbin0", "nephbin1",
  "nephbin2", "nephbin3", "nephbin4", "nephbin5", "nephpm1", "nephpm10", "nephpm25", "opcbin0", "opcbin1",
  "opcbin10", "opcbin11", "opcbin12", "opcbin13", "opcbin14", "opcbin15", "opcbin16", "opcbin17", "opcbin18",
  "opcbin19", "opcbin2", "opcbin20", "opcbin21", "opcbin22", "opcbin23", "opcbin3", "opcbin4", "opcbin5",
  "opcbin6", "opcbin7", "opcbin8", "opcbin9", "opcpm1", "opcpm10", "opcpm25", "opcrh", "opctemp", "geo.ylat",
  "geo.ylon", "met.yrh", "met.ytemp"
)

modPM <- modPM %>%
  select(-one_of(columns_to_delete))

print(modPM)
```

### Meteorology Data

```{r}
# Import meteorology file
metdata <- suppressMessages({
  list.files(path = "./data/Metdata/",
             pattern = "*.csv", 
             full.names = TRUE) %>%
    purrr::map_dfr(read.csv)
})

print(metdata)
# Setting datetime, using correct timezone on East Coast Local time
metdata$date <- as.POSIXct(metdata$valid, format = "%Y-%m-%d %H:%M", tz = "America/New_York")

metdata <- metdata %>%
  rename(wd = drct, ws = sped, original_met_time = valid) %>%
  na.omit(date) %>%
  complete(date = seq(from = min(date), to= max(date),  by = "1 min")) %>%
  fill(c(wd, ws, tmpc)) #fill those new 1 minute interval rows

metdata$ws <- metdata$ws * (1609/3600) #converting to m/s, 1609 meters per mile, 3600 seconds per hr
metdata[,c("station")] <- list(NULL) #getting rid of unnecessary variables
metdata <- unique(metdata) #remove duplicates
print(metdata)
```

## DATA PROCESSING

### Merging Datasets

```{r}
# Time sync and merge met and pollution data
modPM$modPM_date_1min <- lubridate::round_date(modPM$date, unit = "minute")
metdata$met_date_1min <- lubridate::round_date(metdata$date, unit = "minute")

# Merge the two datasets
modPM_met <- dplyr::left_join(modPM, metdata, by = c("modPM_date_1min" = "met_date_1min"))
modPM_met$date <- modPM_met$date.x

# Delete unnecessary columns
modPM_met <- subset(modPM_met, select = -c(date.x, date.y))
```

### Cleaning Up and Save the Work Space

```{r}
# Remove all objects except the final data frame
rm(list = setdiff(ls(), "modPM_met"))
# Save environment to graphableData.RData
save.image("./data/graphableData.RData")
```

### Sanity Checks

```{r}
# # Uncomment to test timeVar plots
# # Feel free to add any additional tests
# library(openair)
# 
# print(modPM_met)
# openair::timeVariation(modPM_met, pollutant = c("ws", "wd", "tmpc"), normalise = TRUE)
```
